Predicting Covid Data

### 2022-12-05 jin_sim.Rmd
Added introduction and peak day of China.

```{r}
data <- read.csv("https://raw.githubusercontent.com/stat400-csu/project-5/main/owid-covid-data.csv")
View(data)
```

@@

# multiply the nominal value with gau_sample for each day
mc_data <- vector()
for (j in 1:n){
  mc_data <- c(mc_data, china$total_cases[j]*gau_sample[,j])
}
mc_data <- matrix(mc_data, nrow=m, ncol=n)  # daily data to obtain 150 series of cases
head(mc_data)
```

@@

@@

## Day of flex (found without the fitted model)

```{r}
flex <- rep(NA, m)

# for each Monte Carlo simulation (m)
for (i in 1:m){
  temp_data <- mc_data[i,]
  num_diff <- 0
  # for each day
  for (j in 1:(n-1)){
    num_diff <- temp_data[j+1] - temp_data[j]
    if (num_diff < 0){
      flex[i] <- j
      break
    }
  }
}

for (i in 1:m){
  temp_data <- mc_data[i,]
  num_diff <- rep(NA,n-1)
  # for each day
  for (j in 1:(n-1)){
    num_diff[j] <- temp_data[j+1] - temp_data[j]
  }
  for (j in 1:(n-1)){
    if (num_diff[j+1] < num_diff[j]){
      flex[i] <- j
      break
    }
  }
}
flex
```

>>>>

```{r}
data <- read.csv("owid-covid-data.csv")
View(data)
```

@@

# multiply the nominal value with gau_sample for each day
mc_data <- vector()
for (j in 1:n){
  mc_data <- c(mc_data, china$total_cases[j]*gau_sample[,j])
}
mc_data <- matrix(mc_data, nrow=m, ncol=n)  # daily data to obtain 150 series of cases
```

```{r}
# multiply the nominal value with gau_sample for each day
mc_data_fa <- vector()
for (j in 1:n){
  mc_data_fa <- c(mc_data_fa, china$total_deaths[j]*gau_sample[,j])
}
mc_data_fa <- matrix(mc_data_fa, nrow=m, ncol=n)  # daily data to obtain 150 series of cases
```

@@

```{r}
flex_day_fa <- rep(NA, m)
for (i in 1:m){
  try <- mc_data_fa[i,]
  model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
  est_try <- summary(model_try)$parameters
  est_try <- est_try[, 1]
  names(est_try) <- NULL
  fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
  erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
  fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
  fn_try_deriv <- D(fn_try_deriv,'x')
  x=seq(1,67,length.out=6700)
  tmp <- eval(fn_try_deriv)
  flex_day_fa[i] <- min(x[tmp < 0])
}

mean(flex_day_fa)
sd(flex_day_fa)
```

@@

```{r}
ggplot() +
  geom_point(aes(x=1:m, y=flex_day_fa), color="red") +
  geom_hline(yintercept=mean(flex_day_fa), color="blue") +
  labs(x = "Number of Monte Carlo simulation", y = "Day of flex, from February 15")
```

## Day of flex (no MC)

```{r}
try <- china$total_cases
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

```{r}
try <- china$total_deaths
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

## superimpose the peak day to the plots

```{r, warning=FALSE}
fn_model1 <- function(x) {4.017e+04 + 4.047e+04 * erf(9.284e-02 * x - 1.732e+00)}
positive +
  stat_function(fun=fn_model1, color="blue") +
  geom_segment(aes(x = 18.65517, y = 0, xend = 18.65517, yend = china$total_cases[19]), color="blue")
fn_model2 <- function(x) {1.571e+03 + 1.658e+03 * erf(6.194e-02 * x - 1.533e+00)}
fatalities +
  stat_function(fun=fn_model2, color="blue") +
  geom_segment(aes(x = 24.74384, y = 0, xend = 24.74384, yend = china$total_deaths[25]), color="blue")
```

### 2022-12-05 main.rmd
Added introduction and peak day of China.

## The MonteCarlo Study

@@

@@

## Available Official Data

Cumulative diagnosed positive cases:

@@

# recreate plots
positive <- ggplot(data = china) +
  geom_point(aes(x = days, y = total_cases), color = "red") +
  labs(x = "Days", y = "China cumulative number of cases")
positive
```

## Available Official Data

Cumulative fatalities:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
fatalities <- ggplot(data = china) +
  geom_point(aes(x = days, y = total_deaths), color = "red") +
  labs(x = "Days", y = "China cumulative number of fatalities") +
  ylim(0,5000)
fatalities
```

@@

Fit of cumulative diagnosed positive cases:

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
model1 <- nls(total_cases ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
summary(model1)$parameters
fn_model1 <- function(x) {4.017e+04 + 4.047e+04 * erf(9.284e-02 * x - 1.732e+00)}
positive + stat_function(fun=fn_model1, color="blue")
```

Fit of cumulative fatalities:

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
model2 <- nls(total_deaths ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
#summary(model2)$parameters
fn_model2 <- function(x) {1.571e+03 + 1.658e+03 * erf(6.194e-02 * x - 1.533e+00)}
fatalities +
  stat_function(fun=fn_model2, color="blue")
```

>>>>

## The Monte Carlo Study

@@

## Ciufolini and Paolozzi's Study

* Mathematical prediction of the time evolution of the COVID-19 pandemic by a Gauss error function and Monte Carlo simulations:
  * The number of cumulative diagnosed positive cases and fatalities in China approximates a gauss error function: $a+b*erf(c*x-d)$.
  * Predicted the potential number of cases in Italy.
  * Predicted the day in which the peak of the number of daily cases in Italy occurs.

@@

## Available Official Data (Positive Cases)

@@

# recreate plots
positive <- ggplot(data = china) +
  geom_point(aes(x = date, y = total_cases), color = "red") +
  labs(x = "Date", y = "China cumulative number of cases")
positive
```

## Available Official Data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
ggplot(data = china) +
  geom_point(aes(x = date, y = total_deaths), color = "red") +
  labs(x = "Date", y = "China cumulative number of fatalities") +
  ylim(0,5000)
```

@@

## Fit by the Gauss Error Function and Peak Day (Positive Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
try <- china$total_cases
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=seq(1,67,length.out=6700)
tmp <- eval(fn_try_deriv)
peak1 <- min(x[tmp < 0])
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
model1 <- nls(total_cases ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
# summary(model1)$parameters
fn_model1 <- function(x) {4.017e+04 + 4.047e+04 * erf(9.284e-02 * x - 1.732e+00)}
ggplot(data = china) +
  geom_point(aes(x = days, y = total_cases), color = "red") +
  labs(x = "Days", y = "China cumulative number of cases") +
  stat_function(fun=fn_model1, color="blue") +
  geom_segment(aes(x = peak1, y = 0, xend = peak1, yend = china$total_cases[ceiling(peak1)]))
```

## Fit by the Gauss Error Function and Peak Day (Fatalities)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
try <- china$total_deaths
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=seq(1,67,length.out=6700)
tmp <- eval(fn_try_deriv)
peak2 <- min(x[tmp < 0])
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
model2 <- nls(total_deaths ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
#summary(model2)$parameters
fn_model2 <- function(x) {1.571e+03 + 1.658e+03 * erf(6.194e-02 * x - 1.533e+00)}
ggplot(data = china) +
  geom_point(aes(x = days, y = total_deaths), color = "red") +
  labs(x = "Days", y = "China cumulative number of fatalities") +
  ylim(0,5000) +
  stat_function(fun=fn_model2, color="blue") +
  geom_segment(aes(x = peak2, y = 0, xend = peak2, yend = china$total_deaths[ceiling(peak2)]))
```
